---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
Load the data:
```{r}
mydata <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Ignoring missing values.
Histogram of the total number of steps taken each day:
```{r}
hist(mydata$steps, main=paste("Histogram of # of steps"), xlab = "# of steps")
```
Mean and median total number of steps taken per day:
```{r}
library(plyr)
ddply(mydata, ~date, summarise, mean=mean(steps, na.rm=TRUE), median=median(steps, na.rm=TRUE))
```

## What is the average daily activity pattern?
Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):
```{r}
library(ggplot2)
aggr <- aggregate(mydata$steps, by=list(mydata$interval), FUN=mean, na.rm=TRUE)
names(aggr) <- c("interval","avg")
ggplot(data=aggr, aes(x=interval, y=avg)) + 
  geom_line() +
    ggtitle("Avg # steps")
```
Interval with the maximum number of steps across all days:
```{r}
aggr[which.max(aggr$avg),]
```

## Imputing missing values
Imputing missing values.
Total number of missing values in the dataset:
```{r}
sum(is.na(mydata$steps))
```
Replace missing values with median for given interval across all days:
```{r}
mydata2 <- ddply(mydata, 
      .(interval), 
      transform, 
      steps=ifelse(is.na(steps), median(steps, na.rm=TRUE), steps))
```

## What is mean total number of steps taken per day?
Ignoring missing values.
Histogram of the total number of steps taken each day:
```{r}
hist(mydata2$steps, main=paste("Histogram of # of steps"), xlab = "# of steps")
```
Mean and median total number of steps taken per day:
```{r}
library(plyr)
ddply(mydata2, ~date, summarise, mean=mean(steps, na.rm=TRUE), median=median(steps, na.rm=TRUE))
```

## Are there differences in activity patterns between weekdays and weekends?
Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis), for weekdays vs weekends: 
```{r}
mydata2$weekday <- weekdays(as.Date(mydata2$date))
mydata2$weekday[mydata2$weekday=="Saturday" | mydata2$weekday=="Sunday"] <- "weekend"
mydata2$weekday[mydata2$weekday!="weekend"] <- "weekday"
mydata2$weekday <- factor(mydata2$weekday, labels=c("weekday","weekend"))

aggr2 <- aggregate(mydata2$steps, by=list(mydata2$interval,mydata2$weekday), FUN=mean)
names(aggr2) <- c("interval","weekday","avg")
ggplot(data=aggr2, aes(x=interval, y=avg)) + 
     geom_line() +
     ggtitle("Avg # steps") +
     facet_grid(weekday ~ .)
```

